\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\setlength{\parindent}{1em}
\setlength{\parskip}{1em}

\begin{document}
	\title{Adellian: Lore}
	\maketitle
	
	\section{Preamble}
	
	No language exists in isolation. This page tries to create the story surrounding Adellians, their customs, culture, religion, law, and so on. The concept is in early work and might change significantly later on.
	
	\tableofcontents
	
	\section{General information}
	
	Adella has various legends that tell the stories of significantly more developed civilization existing long before their own, but then collapsing because of a technological catastrophe. Many of the religious and other ancient texts Adellians consider to be the shards of the knowledge of that ancient civilization.
	
	It is unclear how true these stories are; alternative hypotheses exist that this is merely a way to give more weight to the discoveries of Adellian religious figures of today, or, alternatively, traces of parallel-world contacts happening before our own. At the same time, if this theory is to be confirmed, that does explain a lot of the peculiarities we observe.
	
	Currently, the technological development of Adella, on average, reminds of late XVIII-early XIX century on Earth. However, it would not be correct to draw a direct comparison, as their level of development varies significantly from field to field and does not directly correspond to any known historical era.
	
	The political system is complicated, but also seems to be decentralized almost to utterly medieval extent. The institutes, while full of complicated checks-and-balances, cannot be really considered democratic, and the laws are deeply tied with their tritheistic religion Adellian fully and wholeheartedly believe in their vast majority.
	
	The theoretical medical knowledge is quite advanced, which also influences some of the existing cultural customs; at the same time, there isn't enough technological invention to make use of the full potential of their academy base.
	
	The cultural values seem, in some regards, surprisingly humane for a society of their level of development, and pretty cruel in other respects. A traveller that wishes to pay a visit to Adella must be informed about these cultural differences and careful to not to get in any trouble.
	
\end{document}